\thetitle

### 2021-05-08 paper/paper.tex
results + partly conclusion

A major problem occurring at average reward RL is that the bias values are not uniquely defined
without solving the first set of constraints defined by the error term addends
\citep[see][p.346]{Puterman94,Mahadevan96_SensitiveDiscountOptimalityUnifyingDiscountedAndAverageRewardReinforcementLearning}.
% We could overcome this issue by simply requiring \(\gamma\) to be strictly less than \(1\).
%
Our algorithm, based on the tabular version of \citet{schneckenreither2020average}, does not require
the exact solution for \(V^{\pol}(s)\), but a solution which is offset suffices. Clearly this
observation reduces the required iteration steps tremendously as finding the exact solution,
especially for large discount factors, is tedious. Therefore, we allow to set \(\gamma = 1\), which
induces \(\X_{\gamma}^{\pol}(s) = V^{\pol}(s) + u\), where \(u\) is for unichain MDPs a scalar value
independent of \(s\), i.e.\@ equivalent for all states of the MDP~\cite[p.346]{Puterman94}.
%
%% Nonetheless, we require \(\gamma < 1\) to stabilise the process.
%
If we are interested in correct bias values, i.e. \(\gamma\) is sufficiently close but strictly less
than \(1\), our approach is a tremendous advantage over average reward RL as it reduces the number
of iterative learning steps by requiring only a single constraint per state plus one for the scalar
average reward value. That is, for an MDP with \(N\) states only one more constraint (\(N+1)\) has
to be solved in \ARA{} as compared to (at least) \(2N+1\) nested constraints for average reward RL.
Therefore, it is cheap to compute \(\X_{\gamma}^{\pol}(s)\), while it is rather expensive to find
the correct values of \(V^{\pol}(s)\) directly, especially in an iterative manner as RL is.

@@

\paragraph{Average Reward Adjusted Deep RL Algorithm.} %
The model-free average reward adjusted RL algorithm is based on the tabular version of
\citet{schneckenreither2020average} and depicted in Algorithm~\ref{alg:near}. The algorithm operates
on a target network \(\netTarget{}\) and a worker network \(\netWorker{}\) as in
\citet{mnih2015human}. % The target network reduces fluctuations in the state value estimates and
% therefore reduces the
%
In the action selection process a \(\epsilon\)-sensitive lexicographic order
\(a=(a_{1},\ldots,a_{n}) \lex (b_{1},\ldots,b_{n})=b\) is used. It is defined as \(a \lex b\) if and
only if \(| a_{j} - b_{j} | \leqslant \epsilon\) for all \(j < i\) and
\(|a_{i} - b_{i}| > \epsilon\). Note that the resulting sets of actions may not be disjoint, but
taking the maximum as required in our algorithm is straight-forward and thus cheap to compute. The
first action selection criteria selects the actions which maximise the bias values, the second the
error term.
%
Equation~\ref{eq:constr2} of the average reward RL constraints is used to estimate the average
reward based on the \MS[t]{reformulate avg RL constraints using s' etc.?} currently inferred state
values of two consecutive states and the returned reward. % As the average
% reward is equal among all states estimating the average reward is based on many observations.
This infers a solid average reward prediction of the current policy in few steps only and thus
quickly leads to better policies, as ones with higher gain are preferred.
\citet{schneckenreither2020average} and
\citet{Tadepalli98_ModelbasedAverageRewardReinforcementLearning} find that updating the average
reward by Equation~\ref{eq:constr2} is superior as compared to exponentially smoothing the actual
observed rewards. This makes sense as the later rather evaluates past policies.

@@

For instance, with \(90\%\) utilisation and medium variable demand stream, i.e.\@ 90-(\(\Unif\)), we
can see that \rlt{0.99} performs best (\(576.48\)), followed by \qlt{0.99} (\(586.92\)) and
\BILThree{} (\(595.35\)), as well as \rlt{1.00} (\(602.54\)), where the later two are not
significantly different. In this case \rlt{0.99} also achieves the best service level of \(87.9\%\)
within this group of best performing algorithms. The WIPC costs, as well as the SFTT, show that the
reinforcement learning agents have a higher work-in-process, which is clear as they try to smooth
the workload of demand stream. When looking at the timing cost it becomes even more apparent that
the agents perform well in timing as \rlt{0.99} performs much better (\(471.13\)) than the other two
agents (\(488.01\), \(494.66\)) and all the \BIL{} techniques accumulated more \fgibocName{} costs
(\(507.10\) for the best performing \BIL{}, \BILThree{}).

@@

For \(70\%\) utilization of the bottleneck, i.e. 80-\(\Exp\), \BILThree{} performs best with costs
of \(345.80\), followed by \rlt{0.99} with \(358.41\), and then by \qlt{0.99} and \BILTwo{}
(\(360.70\), \(364.18\)). In terms of timing performance, again \ARA{} outperforms the opponents
with \fgibocName{} costs of \(296.99\), and \BILThree{} with the second highest costs (\(297.17\)).
However, the service level of \BILThree{} is higher with \(93.7\%\) due to the lower shop floor
throughput time (SFTT, \(1.48\) for \BILThree{} versus \(1.73\) for \rlt{0.99}).

% Table~\ref{tbl:res} shows the results, where we evaluate the average reward and discounted
% reinforcement learning algorithm by providing either direct reward feedback, that is the reward
% accumulated in the next period, without keeping track of any orders, as \(\ActShipped\) and
% \(\ActShippedDQN\) respectively. Further both algorithms are evaluated with keeping track of the
% order pool orders, such that each action is responsible for all orders in the order pool. These are
% labelled \(\ActOrdPool\) and \(\ActOrdPoolDQN\). For comparisons reasons we provide
% \(\BILOne-\BILSix\) and immediate release (interval release).

% The results show that at the current time we are facing issues with scaling the algorithm to the
% order release problem. Likely this is due to parameterisation issues, which we are investigating
% currently. However, it can be seen that i) the technique of keeping track of the order pool orders
% leads to better policies and ii) that despite the parameterisation issues average reward
% reinforcement learning finds better solutions than its discounted counterpart. However, the least
% costs are generated by \(\BILThree\), which shows that there is room for improvement for the
% reinforcement learning policies.

% Interesting is also to see, that all reinforcement learning variants are releasing rather late,
% leading to low finished goods inventory costs (FGIC), but high backorder costs (BOC). Furthermore,
% the agents based on the order-pool orders, namely \(\ActOrdPool\) and \(\ActOrdPoolDQN\), are able
% achieve lower work-in-process costs (WIPC) and also lower tardiness, than the other the agents
% rewarded directly on the shipped orders.

% \begin{table}[t!]
%   \centering
%   \def\arraystretch{1.2}          %  1 is the default, change whatever you need
%   \setlength\tabcolsep{4pt}
%   \begin{tabular}{|l||c|c|c|c|c|c|c|c|c|}
%     % BEGIN RECEIVE ORGTBL res2
%     \hline
%     Algorithm/KPI & SUM & BOC & FGIC & WIPC & % \%SL &
%                                                        TARD & \(\sigma\)TARD & SFTT\\
%     \hline
%     % Evaluation 1 & \multicolumn{11}{c|}{} \\
%     \ActShipped    & 1777.74 & 1322.04 & 94.49 & 361.20       & 2.24 & 1.52 & 3.44\\
%     \ActOrdPool    & 1593.03 & 1318.65 &14.01 &260.36         & 2.11 & 1.42 & 2.68\\
%     \ActShippedDQN & 1940.62 & 1570.40 &6.30 &363.92          & 2.35 & 1.59 & 3.95 \\
%     \ActOrdPoolDQN & 1618.88 & 1306.20 & 40.68& 272.00        & 2.08 & 1.38 & 2.58\\
%     \BILOne        & 1736.98 & 1510.42  &  0.00  & 226.56     & 2.04 & 1.26 & 2.16 \\
%     \BILTwo        & 1081.79 & 769.86 & 86.89   & 225.03      & 1.87 & 1.18 & 2.15\\
%     \BILThree      &  \textbf{922.36} & 362.59 &  334.47  & 225.30     & 1.74 & 1.07 & 2.15\\
%     \BILFour       & 1062.68 & 155.72  & 681.03   & 225.93    & 1.63 & 0.99 & 2.15\\
%     \BILFive       & 1380.04 & 65.45  & 1088.94    & 225.64   & 1.67 & 1.02 & 2.15\\
%     \BILSix        & 1777.77 & 25.04 &  1529.77  & 222.96     & 1.61 & 0.85 & 2.13\\
%     \imre          & 2209.97 & 12.08 &1974.72 & 223.16        & 1.58 & 0.63 & 2.14 \\
%     \hline
%     % END RECEIVE ORGTBL res2
%   \end{tabular}
%         %         \egroup

%   \caption{\label{tbl:res}Evaluation results where the monetary units are in \(k\)-values.}

%         %         \end{figure}
% \end{table}

@@

>>>>

A major problem occurring at average reward RL is that the bias values are not uniquely defined
without solving the first set of constraints defined by the error term addends
\citep[see][p.346]{Puterman94,Mahadevan96_SensitiveDiscountOptimalityUnifyingDiscountedAndAverageRewardReinforcementLearning}.
% We could overcome this issue by simply requiring \(\gamma\) to be strictly less than \(1\).
%
Our average reward adjusted learning algorithm (\ARA{}), based on the tabular version of
\citet{schneckenreither2020average}, does not require the exact solution for \(V^{\pol}(s)\), but a
solution which is offset suffices. Clearly this observation reduces the required iteration steps
tremendously as finding the exact solution, especially for large discount factors, is tedious.
Therefore, we allow to set \(\gamma = 1\), which induces
\(\X_{\gamma}^{\pol}(s) = V^{\pol}(s) + u\), where \(u\) is for unichain MDPs a scalar value
independent of \(s\), i.e.\@ equivalent for all states of the MDP~\cite[p.346]{Puterman94}.
%
%% Nonetheless, we require \(\gamma < 1\) to stabilise the process.
%
If we are interested in correct bias values, i.e. \(\gamma\) is sufficiently close but strictly less
than \(1\), our approach is a tremendous advantage over average reward RL as it reduces the number
of iterative learning steps by requiring only a single constraint per state plus one for the scalar
average reward value. That is, for an MDP with \(N\) states only one more constraint (\(N+1)\) has
to be solved in \ARA{} as compared to (at least) \(2N+1\) nested constraints for average reward RL.
Therefore, it is cheap to compute \(\X_{\gamma}^{\pol}(s)\), while it is rather expensive to find
the correct values of \(V^{\pol}(s)\) directly, especially in an iterative manner as RL is.

@@

\paragraph{Average Reward Adjusted Deep RL Algorithm.} %
The model-free average reward adjusted RL algorithm is based on the tabular version of
\citet{schneckenreither2020average} and depicted in Algorithm~\ref{alg:near}. The algorithm operates
on a target network \(\netTarget{}\) and a worker network \(\netWorker{}\) as in
\citet{mnih2015human}. % The target network reduces fluctuations in the state value estimates and
% therefore reduces the
%
In the action selection process a \(\epsilon\)-sensitive lexicographic order
\(a=(a_{1},\ldots,a_{n}) \lex (b_{1},\ldots,b_{n})=b\) is used. It is defined as \(a \lex b\) if and
only if \(| a_{j} - b_{j} | \leqslant \epsilon\) for all \(j < i\) and
\(|a_{i} - b_{i}| > \epsilon\). Note that the resulting sets of actions may not be disjoint, but
taking the maximum as required in our algorithm is straight-forward and thus cheap to compute. The
first action selection criteria selects the actions which maximise the bias values, the second the
error term.
%
Equation~\ref{eq:constr2} of the average reward RL constraints is used to estimate the average
reward based on the currently inferred state values of two consecutive states and the returned
reward. % As the average
% reward is equal among all states estimating the average reward is based on many observations.
This infers a solid average reward prediction of the current policy in few steps only and thus
quickly leads to better policies, as ones with higher gain are preferred.
\citet{schneckenreither2020average} and
\citet{Tadepalli98_ModelbasedAverageRewardReinforcementLearning} find that updating the average
reward by Equation~\ref{eq:constr2} is superior as compared to exponentially smoothing the actual
observed rewards. This makes sense as the later rather evaluates past policies.

@@

For instance, with \(90\%\) utilisation and medium variable demand stream, i.e.\@ 90-(\(\Unif\)), we
can see that \rlt{0.99} performs best (\(576.48\)), followed by \qlt{0.99} (\(586.92\)) and
\BILThree{} (\(595.35\)), as well as \rlt{1.00} (\(602.54\)), where the later two are not
significantly different. In this case \rlt{0.99} also achieves the best service level of \(87.9\%\)
within this group of best performing algorithms. The WIPC costs, as well as the SFTT, show that the
reinforcement learning agents have a higher work in process, which is clear as they try to smooth
the workload of demand stream. When looking at the timing cost it becomes even more apparent that
the agents perform well in timing as \rlt{0.99} performs much better (\(471.13\)) than the other two
agents (\(488.01\), \(494.66\)) and all the \BIL{} techniques accumulated more \fgibocName{} costs
(\(507.10\) for the best performing \BIL{}, \BILThree{}).

@@

For \(70\%\) utilization of the bottleneck, i.e. 80-\(\Exp\), \BILThree{} performs best with costs
of \(345.80\), followed by \rlt{0.99} with \(358.41\), and then by \qlt{0.99} and \BILTwo{}
(\(360.70\), \(364.18\)). In terms of timing performance, again \ARA{} outperforms the opponents
with \fgibocName{} costs of \(296.99\), and \BILThree{} with the second highest costs (\(297.17\)).
However, the service level of \BILThree{} is higher with \(93.7\%\) due to the lower shop floor
throughput time (SFTT, \(1.48\) periods for \BILThree{} versus \(1.73\) periods for \rlt{0.99}).

In case of 80-\(\Exp\) \BILThree{} performs best with costs of \(521.40\), but is not significantly
better than \rlt{1.00} with \(528.54\). Also \rlt{0.99} performs quite well (\(529.00\)). The
service level of all three algorithm are good, but the order is reversed: \rlt{0.99} with
\(87.6\%\), \rlt{1.00} with \(86.6\%\) and \BILThree{} with \(85.8\%\). Again the SFTT of the RL
agents are slightly higher, i.e. \(0.13\)-\(0.20\) periods, than the ones of the \BIL{} release
procedures. The other three \BIL{} methods, \BILOne{}, \BILTwo{} and \BILFour{}, are performing
rather bad (\(1,338.44\), \(679.24\) and \(591.82\) respectively). Interesting to see is that in
this scenario \rlt{0.99} and \rlt{1.00} handle the due date performance better than \qlt{0.99}, as
can be seen in the backorder costs (BOC). Here the former two accumulate costs of \(185.69\) and
\(198.46\), while \qlt{0.99} faces costs of \(225.15\). This results in timing costs (\fgibocName{})
of \(439.16\) for \rlt{0.99} and \(441.37\) for \rlt{1.00}, which also means a better timing cost
than \BILThree{} with \(442.37\). For this scenario the actual measured average reward (average
costs) of about \(75.5\) for both \ARA{} variants and learned average rewards of \(70.53\) for
\rlt{0.99} and \(72.94\) for \rlt{1.00} are a little bit offset. This probably results from the
higher variability in the system.

Finally, for the case of 90-\(\Exp\) \rlt{1.00} performs much better than the other variants with
cots of \(986.32\). It is followed by \rlt{0.99} with costs of \(1,072.06\) and rather far, but not
significantly, afterwards follows \BILFour{} (\(1,104.29\)). As before the shopfloor throughput time
(SFTT) is higher for the RL variants, and with that of course also the work in process costs (WIPC).
The backorder costs of \rlt{1.00} (\(539.80\)) are the lowest of all tested methods. Like for the
overall costs also in terms of timing costs (\fgibocName{}) \rlt{1.00} performs best (\(825.87\)),
followed by \rlt{0.99} (\(914.03\)) and \BILFour{} (\(957.21\)).
%
\BILThree{} and \ql{0.99} perform quite similar in the sense of accumulated costs, \(1,175.88\) and
\(1,179.46\) respectively. But the other measures show that \ql{0.99} has a better timing
performance, thus also a slightly high service level with \(71.5\%\) as compared to \BILThree{} with
\(70.6\%\). The best service level is achieved by \BILFour{} with \(82.5\%\), before \rlt{1.00}
\(79.6\%\), but with a high cost as the difference of SUM is \(117.97\).
%
Interesting to see is that the measured average reward and the learned average reward are very
different for both \ARA{} variants (about \(30\)-\(40\)). This is probably due to the very high
variability in the system, but might also be an indicator that the agents need more time for
learning. However, due to the high computational complexity we encountered we leave this to future
work.

Overall we can see that the RL agents, and especially the established \ARA{} algorithm performs very
well in systems with high variability. In case of the less variable systems the results reveal that
there is still potential for the improvement of reinforcement learning variants.

@@

This paper is one of the first papers that use reinforcement learning for setting lead times to
release orders into the production system. We present a novel average reward adjusted reinforcement
learning algorithm that uses an artificial neural network to approximate the state-action value
functions. We add several extensions to this algorithm, including n-step learning, and a special
experience replay memory that stores its experiences according to the actions takes to prevent
catastrophic forgetting. Additionally we invent \textit{independent agents} to reduce the action
space by independently choosing a subset of possible actions, while operating on the same state.
Additionally we adapt the deep Q-Learning algorithm of \citet{mnih2015human} to be able to present a
fair comparison between the algorithms, where we also use several backward infinite loading
techniques.

The results suggest that for high variablity the RL learning agents
